Exploration of Adversarial Training

### 2025-05-26 AIAA2290/report/main.tex
Update report of AIAA2290
Update ...

\usepackage{algorithm,algpseudocode,float}
\usepackage{lipsum}
\usepackage{longtable}
\usepackage{multirow}

@@

\begin{equation}
\theta^*=\arg\min_\theta\left\{
  \mathbb{E}_{(\mathbf{x}, \mathbf{y})\in \mathcal{D}}
  \left[
      \max_{\lVert\mathbf{\delta}\rVert\leq \epsilon}
          \left\{L(f_\theta(\mathbf{x}+\delta), y)\right\}
  \right]
\right\}
\tag{2:1}
\label{formula:adversarial_target}
\end{equation}

@@

\begin{equation}
\theta^{*} = \arg\min_{\theta} ; \mathbb{E}{(\mathbf{x},\mathbf{y}) \sim \mathcal{D}} \Biggl[\underbrace{\max_{{\lVert \mathbf{\delta} \rVert \le \epsilon}} L\bigl(f_{\theta}(\mathbf{x} + \mathbf{\delta}), \mathbf{y}\bigr)}_{\text{worst-case loss}}\Biggr]
\label{eq:adv_training_objective}
\end{equation}

Here, the inner maximization identifies the most adversarial perturbation $\mathbf{\delta}$ within the $\epsilon$-ball around the input $\mathbf{x}$, while the outer minimization optimizes the model parameters to minimize the expected worst-case loss over the data distribution $\mathcal{D}$.

@@

$$
\begin{aligned}
\mathbf{\delta}_{0} &= \mathbf{0}, \\
\mathbf{\delta}_{t+1} &= \Pi_{{\lVert \mathbf{\delta_t} \rVert \le \epsilon}} \Bigl(\mathbf{\delta}_{t} + \alpha \cdot \mathrm{sign}\bigl(\nabla_{\mathbf{x}} L(f_{\theta}(\mathbf{x} + \mathbf{\delta}_{t}), \mathbf{y})\bigr)\Bigr),
\end{aligned}
$$

@@

After $k$ iterations, the adversarial example is given by $\mathbf{x}^{\mathrm{adv}} = \mathbf{x} + \mathbf{\delta}_{k}$. Incorporating these examples into the training process yields the PGD adversarial training algorithm

@@

  \For {$n=1$ to $N$}
    \State {divide $\mathcal{D}$ into list of mini-batches $\mathcal{C}=\{(\mathrm{X}, \mathbf{y})\}$}
    \For {$(\mathrm{X}, \mathbf{y})\in \mathcal{C}$} 
      \State {$\delta\leftarrow$ zero matrix with shape of $X$}
      \For {$t\leftarrow 1$ to $k$}
        \State {execute backpropagation}
        \State $g_a\leftarrow \nabla_{\mathrm{X}+\delta}(L(f_{\theta}(\mathrm{X}+\delta), \mathbf{y}))$
        \State $\delta\leftarrow \mathrm{clamp}(\delta + \frac{\alpha}{\left\lVert g_a \right\rVert}g_a, -\epsilon, \epsilon)$
      \EndFor
      \State $g_\theta\leftarrow \nabla_{\theta}(L(f_{\theta}(\mathrm{X}+\delta), \mathbf{y}))$
      \State $\theta \leftarrow \theta - \tau g_\theta$
    \EndFor
  \EndFor
\end{algorithmic}
\end{breakablealgorithm}

@@

\begin{breakablealgorithm}
\caption{YOPO training}
\begin{algorithmic}[1]
  \Require training set $\mathcal{D}$, learning rate $\tau$, training epochs $N$, hyperparameters $\alpha, \epsilon, m, n$, model $f_\theta$ and loss function $L$ 
  \State divide the model parameters $\theta$ into first layer $\theta_0$ and the other layers $\theta'$
  \For {$i\leftarrow 1$ to $N$}
    \State {divide $\mathcal{D}$ into list of mini-batches $\mathcal{C}=\{(\mathrm{X}, \mathbf{y})\}$}
    \For {$(\mathrm{X}, \mathbf{y})\in \mathcal{C}$}
      \State $\delta_{0, n}\leftarrow \text{zero matrix with shape of } \mathrm{X}$
      \State $g_\theta\leftarrow \text{zero matrix with shape of } \mathrm{\theta}$
      \For {$j=1$ to $m$}
        \State execute entire backpropagation
        \State $\delta_{j,0}\leftarrow \delta_{j-1,n}$
        \State {$p\leftarrow \nabla_{f_{\theta_0}(\mathrm{X}+\delta_{j,0})}\left(
          L(f_\theta(\mathrm{X}+\delta_{j,0}), \mathbf{y})
        \right)$}
        \For {$k=1$ to $n$}
          \State $g_a\leftarrow
            p\cdot \nabla_{\mathrm{X}+\delta}(f_{\theta_0}(\mathrm{X}+\delta_{j,k-1}))$
          \State $\delta_{j,k}\leftarrow \mathrm{clamp}\left(\delta_{j,k-1}+\frac{\alpha\cdot g_a}{\left\lVert g_a \right\rVert}, -\epsilon, \epsilon\right)$
        \EndFor
        \State $g_\theta\leftarrow g_\theta + \nabla_{\theta}(L(f_{\theta}(\mathrm{X}+\delta_{j,n}), \mathbf{y}))$
        \Comment this gradient can be calculated in backpropagation
      \EndFor
      \State $\theta\leftarrow \theta-\tau\cdot g_\theta$
    \EndFor
  \EndFor
\end{algorithmic}
\end{breakablealgorithm}

@@

\begin{breakablealgorithm}
\caption{FreeLB training}
\begin{algorithmic}[1]
  \Require training set $\mathcal{D}$, learning rate $\tau$, training epochs $N$, hyperparameters $\alpha, \epsilon, k$, model $f_\theta$ and loss function $L$ .
  \For {$i\leftarrow 1$ to $N$}
    \State {divide $\mathcal{D}$ into list of mini-batches $\mathcal{C}=\{(\mathrm{X}, \mathbf{y})\}$}
    \For {$(\mathrm{X}, \mathbf{y})\in \mathcal{C}$}
      \State $\delta_0\leftarrow \textbf{random}\text{ matrix with shape of } \mathrm{X}$
      \State $g_\theta\leftarrow \text{zero matrix with shape of } \theta$
      \For {$i\leftarrow 1$ to $k$}
        \State {execute backpropagation}
        \State $g_\theta\leftarrow g_\theta + \frac{1}{k}\nabla_{\theta}(L(f_\theta(\mathrm{X}+\delta_{i-1}), \mathbf{y}))$
        \State $g_a\leftarrow \nabla_{\mathrm{X}+\delta}(L(f_\theta(\mathrm{X}+\delta_{i-1}), \mathbf{y}))$
        \State $\delta_i\leftarrow \mathrm{clamp}\left(\delta_{i-1}+\frac{\alpha}{\left\lVert g_a \right\rVert}g_a, -\epsilon, \epsilon\right)$
      \EndFor
      \State {$\theta\leftarrow\theta-\tau g_\theta$}
    \EndFor
  \EndFor
\end{algorithmic}
\end{breakablealgorithm}

@@

$$
\begin{aligned}
  &\mathcal{L}(\theta)=\frac{1}{n}\sum_{(\mathbf{x}, \mathbf{y})\in\mathcal{D}} L(f_\theta(\mathbf{x}), \mathbf{y}) \\
  &\mathcal{R}(\theta)=\frac{1}{n}\sum_{(\mathbf{x}, \mathbf{y})\in\mathcal{D}} \max_{\left\Vert \tilde{\mathbf{x}}-\mathbf{x}\right\Vert\leq \epsilon}\left\{l\left(f_\theta(\tilde{\mathbf{x}}), f_\theta(\mathbf{x})\right)\right\}
\end{aligned}
$$

@@

$$
\begin{aligned}
l(f, g)&=\mathrm{KL}(f\Vert g) + \mathrm{KL}(f\Vert g) \\
&=\sum_{x\in \mathcal{X}} f(x)\log\frac{f(x)}{g(x)} + g(x)\log\frac{g(x)}{f(x)} \\
&=\sum_{x\in \mathcal{X}} (f(x) - g(x))\log\frac{f(x)}{g(x)}
\end{aligned}
$$

@@

$$
\mathrm{Breg}_\mathcal{D}(\theta_1, \theta_2)=\frac{1}{\left\vert\mathcal{D}\right\vert}
  \sum_{(\mathbf{x}, \mathbf{y})\in\mathcal{D}}l\left(f_{\theta_1}(\mathbf{x}), f_{\theta_2}(\mathbf{x})\right)
$$

@@

$$
g_\theta(\mathcal{D}, \delta)=
\begin{bmatrix}
  \nabla_{\delta_1}l(f_\theta(\mathbf{x}_1), f_\theta(\mathbf{x}_1+\delta_1)) \\
  \nabla_{\delta_2}l(f_\theta(\mathbf{x}_2), f_\theta(\mathbf{x}_2+\delta_2)) \\
  \vdots \\
  \nabla_{\delta_{|\mathcal{D}|}}l\left(f_\theta\left(\mathbf{x}_{|\mathcal{D}|}\right), f_\theta\left(\mathbf{x}_{|\mathcal{D}|}+\delta_{|\mathcal{D}|}\right)\right)
\end{bmatrix}
$$

@@

  \State $\tilde{\theta_1}\leftarrow\theta_0$ 
  \For {$t\leftarrow 1$ to $N$}
    \State $\overline{\theta}\leftarrow \theta_{t-1}$
    \State {divide $\mathcal{D}$ into mini-batches $\mathcal{C}$}
    \For {$(\mathrm{X}, \mathbf{y})\in \mathcal{C}$}
      \State initialize $\delta$ with guassian distribution
      \State $\forall \delta _i\sim N(0,\sigma^2I)$
      \For {$i=1$ to $k$}
        \State $g_i\leftarrow \frac{g_{\overline{\theta}}\left((\mathrm{X},\mathbf{y}), \delta\right)}{\left\Vert g_{\overline{\theta}}((\mathrm{X},\mathbf{y}), \delta) \right\Vert}$
        \State $\delta\leftarrow \mathrm{clamp}(\delta + \alpha g_i, -\epsilon, \epsilon)$
      \EndFor
      \State $\overline{\theta}\leftarrow \mathrm{Adam}\left(\overline{\theta}, (\mathrm{X}+\delta, \mathbf{y})\right)$
    \EndFor
    \State $\theta_t\leftarrow \overline{\theta}$
    \State $\tilde{\theta}_{t+1}\leftarrow (1-\beta)\overline{\theta}+\beta\tilde{\theta}_t$
  \EndFor
  \State the output is $\theta_N$
\end{algorithmic}
\end{breakablealgorithm}

@@

>>>>

\usepackage{algorithm,algpseudocode,float}
\usepackage{lipsum}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{graphicx}

@@

\begin{equation}
\theta^*=\arg\min_\theta\left\{
  \mathbb{E}_{(\bf{x}, \bf{y})\in \mathcal{D}}
  \left[
      \max_{\lVert\bf{\delta}\rVert\leq \epsilon}
          \left\{L(f_\theta(\bf{x}+\delta), y)\right\}
  \right]
\right\}
\tag{2:1}
\label{formula:adversarial_target}
\end{equation}

@@

\begin{equation}
\theta^{*} = \arg\min_{\theta} ; \mathbb{E}{(\bf{x},\bf{y}) \sim \mathcal{D}} \Biggl[\underbrace{\max_{{\lVert \bf{\delta} \rVert \le \epsilon}} L\bigl(f_{\theta}(\bf{x} + \bf{\delta}), \bf{y}\bigr)}_{\text{worst-case loss}}\Biggr]
\label{eq:adv_training_objective}
\end{equation}

Here, the inner maximization identifies the most adversarial perturbation $\bf{\delta}$ within the $\epsilon$-ball around the input $\bf{x}$, while the outer minimization optimizes the model parameters to minimize the expected worst-case loss over the data distribution $\mathcal{D}$.

@@

$$
\begin{aligned}
\bf{\delta}_{0} &= \bf{0}, \\
\bf{\delta}_{t+1} &= \Pi_{{\lVert \bf{\delta_t} \rVert \le \epsilon}} \Bigl(\bf{\delta}_{t} + \alpha \cdot \mathrm{sign}\bigl(\nabla_{\bf{x}} L(f_{\theta}(\bf{x} + \bf{\delta}_{t}), \bf{y})\bigr)\Bigr),
\end{aligned}
$$

@@

After $k$ iterations, the adversarial example is given by $\bf{x}^{\mathrm{adv}} = \bf{x} + \bf{\delta}_{k}$. Incorporating these examples into the training process yields the PGD adversarial training algorithm

@@

  \For {$n=1$ to $N$}
    \State {divide $\mathcal{D}$ into list of mini-batches $\mathcal{C}=\{(\mathrm{X}, \bf{y})\}$}
    \For {$(\mathrm{X}, \bf{y})\in \mathcal{C}$} 
      \State {$\delta\leftarrow$ zero matrix with shape of $X$}
      \For {$t\leftarrow 1$ to $k$}
        \State {execute backpropagation}
        \State $g_a\leftarrow \nabla_{\mathrm{X}+\delta}(L(f_{\theta}(\mathrm{X}+\delta), \bf{y}))$
        \State $\delta\leftarrow \mathrm{clamp}(\delta + \frac{\alpha}{\left\lVert g_a \right\rVert}g_a, -\epsilon, \epsilon)$
      \EndFor
      \State $g_\theta\leftarrow \nabla_{\theta}(L(f_{\theta}(\mathrm{X}+\delta), \bf{y}))$
      \State $\theta \leftarrow \theta - \tau g_\theta$
    \EndFor
  \EndFor
\end{algorithmic}
\end{breakablealgorithm}

@@

\begin{breakablealgorithm}
\caption{YOPO training}
\begin{algorithmic}[1]
  \Require training set $\mathcal{D}$, learning rate $\tau$, training epochs $N$, hyperparameters $\alpha, \epsilon, m, n$, model $f_\theta$ and loss function $L$ 
  \State divide the model parameters $\theta$ into first layer $\theta_0$ and the other layers $\theta'$
  \For {$i\leftarrow 1$ to $N$}
    \State {divide $\mathcal{D}$ into list of mini-batches $\mathcal{C}=\{(\mathrm{X}, \bf{y})\}$}
    \For {$(\mathrm{X}, \bf{y})\in \mathcal{C}$}
      \State $\delta_{0, n}\leftarrow \text{zero matrix with shape of } \mathrm{X}$
      \State $g_\theta\leftarrow \text{zero matrix with shape of } \mathrm{\theta}$
      \For {$j=1$ to $m$}
        \State execute entire backpropagation
        \State $\delta_{j,0}\leftarrow \delta_{j-1,n}$
        \State {$p\leftarrow \nabla_{f_{\theta_0}(\mathrm{X}+\delta_{j,0})}\left(
          L(f_\theta(\mathrm{X}+\delta_{j,0}), \bf{y})
        \right)$}
        \For {$k=1$ to $n$}
          \State $g_a\leftarrow
            p\cdot \nabla_{\mathrm{X}+\delta}(f_{\theta_0}(\mathrm{X}+\delta_{j,k-1}))$
          \State $\delta_{j,k}\leftarrow \mathrm{clamp}\left(\delta_{j,k-1}+\frac{\alpha\cdot g_a}{\left\lVert g_a \right\rVert}, -\epsilon, \epsilon\right)$
        \EndFor
        \State $g_\theta\leftarrow g_\theta + \nabla_{\theta}(L(f_{\theta}(\mathrm{X}+\delta_{j,n}), \bf{y}))$
        \Comment this gradient can be calculated in backpropagation
      \EndFor
      \State $\theta\leftarrow \theta-\tau\cdot g_\theta$
    \EndFor
  \EndFor
\end{algorithmic}
\end{breakablealgorithm}

@@

\begin{breakablealgorithm}
\caption{FreeLB training}
\begin{algorithmic}[1]
  \Require training set $\mathcal{D}$, learning rate $\tau$, training epochs $N$, hyperparameters $\alpha, \epsilon, k$, model $f_\theta$ and loss function $L$ .
  \For {$i\leftarrow 1$ to $N$}
    \State {divide $\mathcal{D}$ into list of mini-batches $\mathcal{C}=\{(\mathrm{X}, \bf{y})\}$}
    \For {$(\mathrm{X}, \bf{y})\in \mathcal{C}$}
      \State $\delta_0\leftarrow \textbf{random}\text{ matrix with shape of } \mathrm{X}$
      \State $g_\theta\leftarrow \text{zero matrix with shape of } \theta$
      \For {$i\leftarrow 1$ to $k$}
        \State {execute backpropagation}
        \State $g_\theta\leftarrow g_\theta + \frac{1}{k}\nabla_{\theta}(L(f_\theta(\mathrm{X}+\delta_{i-1}), \bf{y}))$
        \State $g_a\leftarrow \nabla_{\mathrm{X}+\delta}(L(f_\theta(\mathrm{X}+\delta_{i-1}), \bf{y}))$
        \State $\delta_i\leftarrow \mathrm{clamp}\left(\delta_{i-1}+\frac{\alpha}{\left\lVert g_a \right\rVert}g_a, -\epsilon, \epsilon\right)$
      \EndFor
      \State {$\theta\leftarrow\theta-\tau g_\theta$}
    \EndFor
  \EndFor
\end{algorithmic}
\end{breakablealgorithm}

@@

$$
\begin{aligned}
  &\mathcal{L}(\theta)=\frac{1}{n}\sum_{(\bf{x}, \bf{y})\in\mathcal{D}} L(f_\theta(\bf{x}), \bf{y}) \\
  &\mathcal{R}(\theta)=\frac{1}{n}\sum_{(\bf{x}, \bf{y})\in\mathcal{D}} \max_{\left\Vert \tilde{\bf{x}}-\bf{x}\right\Vert\leq \epsilon}\left\{l\left(f_\theta(\tilde{\bf{x}}), f_\theta(\bf{x})\right)\right\}
\end{aligned}
$$

@@

$$
\begin{aligned}
l(f, g)&=\mathrm{KL}(f\Vert g) + \mathrm{KL}(g\Vert f) \\
&=\sum_{x\in \mathcal{X}} f(x)\log\frac{f(x)}{g(x)} + g(x)\log\frac{g(x)}{f(x)} \\
&=\sum_{x\in \mathcal{X}} (f(x) - g(x))\log\frac{f(x)}{g(x)}
\end{aligned}
$$

@@

$$
\mathrm{Breg}_\mathcal{D}(\theta_1, \theta_2)=\frac{1}{\left\vert\mathcal{D}\right\vert}
  \sum_{(\bf{x}, \bf{y})\in\mathcal{D}}l\left(f_{\theta_1}(\bf{x}), f_{\theta_2}(\bf{x})\right)
$$

@@

$$
g_\theta(\mathcal{D}, \delta)=
\begin{bmatrix}
  \nabla_{\delta_1}l(f_\theta(\bf{x}_1), f_\theta(\bf{x}_1+\delta_1)) \\
  \nabla_{\delta_2}l(f_\theta(\bf{x}_2), f_\theta(\bf{x}_2+\delta_2)) \\
  \vdots \\
  \nabla_{\delta_{|\mathcal{D}|}}l\left(f_\theta\left(\bf{x}_{|\mathcal{D}|}\right), f_\theta\left(\bf{x}_{|\mathcal{D}|}+\delta_{|\mathcal{D}|}\right)\right)
\end{bmatrix}
$$

@@

  \State $\tilde{\theta_1}\leftarrow\theta_0$ 
  \For {$t\leftarrow 1$ to $N$}
    \State $\overline{\theta}\leftarrow \theta_{t-1}$
    \State {divide $\mathcal{D}$ into mini-batches $\mathcal{C}$}
    \For {$(\mathrm{X}, \bf{y})\in \mathcal{C}$}
      \State initialize $\delta$ with guassian distribution
      \State $\forall \delta _i\sim N(0,\sigma^2I)$
      \For {$i=1$ to $k$}
        \State $g_i\leftarrow \frac{g_{\overline{\theta}}\left((\mathrm{X},\bf{y}), \delta\right)}{\left\Vert g_{\overline{\theta}}((\mathrm{X},\bf{y}), \delta) \right\Vert}$
        \State $\delta\leftarrow \mathrm{clamp}(\delta + \alpha g_i, -\epsilon, \epsilon)$
      \EndFor
      \State $\overline{\theta}\leftarrow \mathrm{Adam}\left(\overline{\theta}, (\mathrm{X}+\delta, \bf{y})\right)$
    \EndFor
    \State $\theta_t\leftarrow \overline{\theta}$
    \State $\tilde{\theta}_{t+1}\leftarrow (1-\beta)\overline{\theta}+\beta\tilde{\theta}_t$
  \EndFor
  \State the output is $\theta_N$
\end{algorithmic}
\end{breakablealgorithm}

@@

The record of training process is also provided :

\begin{figure}[H]
\centering
\caption{Test With Original Samples}
\includegraphics[width=0.45\textwidth]{figures/plot-Original.png}
\end{figure}

\begin{figure}[H]
\centering
\caption{Test With PGD-10}
\includegraphics[width=0.45\textwidth]{figures/plot-PGD-10 Attack.png}
\end{figure}

\begin{figure}[H]
\centering
\caption{Test With PGD-40}
\includegraphics[width=0.45\textwidth]{figures/plot-PGD-40 Attack.png}
\end{figure}

\begin{figure}[H]
\centering
\caption{On Training set}
\includegraphics[width=0.45\textwidth]{figures/plot-Training.png}
\end{figure}
